Disturbance-aware minimum-time planning strategies for motorsport vehicles with probabilistic safety certificates

### 2025-06-04 performance_comparison.tex
[mg] subsec 5.4

\subsection{Open-loop vs Closed-loop evaluation}
We now compare the open-loop planning with the closed-loop approach. In particular, we consider open-loop and closed-loop trajectories computed using both the robustified TLC and FLC, with $\ga^\textrm{TLC}=\ga^\textrm{FLC}=1.28$. As before, the open-loop planning is carried out with a prediction horizon of $H=4$ steps.

For the TLC, the back-off term $\be^{TLC}$ in~\eqref{eq:TLC} admits a clear geometrical interpretation: it corresponds to the amount by which the track boundaries are tightened to meet the probabilistic constraint, or, equivalently, to a virtual extension of the vehicle's width. According to the latter interpretation, Figure~\ref{fig:olcl_traj_bands} shows two highlighted bands---red for the open-loop case and cyan for the closed-loop case---corresponding to the area swept by the back-off-extended-vehicle. Ultimately, the TLC enforces precisely that these bands remain entirely within the actual track limits. The comparison between the two regions illustrates how the additional conservativeness of the open-loop approach stems from larger back-off terms, i.e., a wider band that must stay within the track boundaries. The closed-loop method, on the other hand, benefits from the LQR-based feedback, which allows it to maintain smaller back-off terms and thus a narrower band, requiring a reduced safety margin from the track limits.

\begin{figure}
	\centering
	\includegraphics{Fig/olcl_traj_bands.pdf}
	\caption{Lorem ipsum.}
	\label{fig:olcl_traj_bands}
\end{figure}

\subsection{Simulations with noise realization}
This analysis aims to provide empirical validation of the robust trajectories obtained using the methods proposed in this work. For conciseness, the comparison is limited to a nominal trajectory and a trajectory computed using the closed-loop method with only the track limit constraint robustified.
An LQR controller, designed according to the procedure outlined in Section~\ref{sec:LQR}, is implemented to stabilize the nominal trajectory. In contrast, the robust trajectory is tracked using the controller directly obtained from the optimization process.

>>>>

\subsection{Open-loop vs closed-loop robust planning}
We now compare the open-loop planning with the closed-loop approach. In particular, we consider open-loop and closed-loop trajectories computed using both the robustified TLC and FLC, with $\ga^\textrm{TLC}=\ga^\textrm{FLC}=1.28$. As before, the open-loop planning is carried out with a prediction horizon of $H=4$ steps.

For the TLC, the back-off term $\be^{TLC}$ in~\eqref{eq:TLC} admits a clear geometrical interpretation: it corresponds to the amount by which the track boundaries are tightened to meet the probabilistic constraint, or, equivalently, to a virtual extension of the vehicle's width. According to the latter interpretation, Figure~\ref{fig:olcl_traj_bands} shows two highlighted bands---red for the open-loop case and cyan for the closed-loop case---corresponding to the area swept by the back-off-extended-vehicle. Ultimately, the TLC enforces precisely that these bands remain entirely within the actual track limits. The comparison between the two regions illustrates how the additional conservativeness of the open-loop approach stems from larger back-off terms, i.e., a wider band that must stay within the track boundaries. The closed-loop method, on the other hand, benefits from the LQR-based feedback, which allows it to maintain smaller back-off terms and thus a narrower band, requiring a reduced safety margin from the track limits.
\begin{figure}
	\centering
	\includegraphics{Fig/olcl_traj_bands.pdf}
	\caption{Lorem ipsum.}
	\label{fig:olcl_traj_bands}
\end{figure}

\subsection{Simulations with noise realization}
Finally, we provide an empirical validation of the closed-loop robustified planning by simulating the evolution of the controlled vehicle with random initial conditions and including noise realizations. The random values for the initial state and the noise follow Gaussian distributions with covariance matrices $\bP_0$ and $\bQ$, respectively, as introduced in~\eqref{eq:dP}. This analysis cannot be performed for the open-loop approach, as it does not compute a stabilizing controller---resulting in trajectories that would significantly deviate from the track limits beyond the prediction horizon. For visual clarity, the comparison is limited to trajectories with the robustified TLC, but analogous considerations apply to the robustified FLC as well.

The left panel of Figure~\ref{fig:traj_strings} shows random evolutions (red lines) for an LQR-controlled vehicle stabilized to follow the nominal optimal trajectory (black line). No robust constraints are introduced, and the controller implements the feedback policy from step 2 of Subsection~\ref{sec:LQR}. The right panel, instead, shows the random evolutions (red lines) for the robustified LQR-based controller obtained in the step 3 of the closed-loop approach described in Subsection~\ref{sec:CLrobustified}. The corresponding robustified mean trajectory is shown in black.
The comparison between the panels shows that the robustified controller is significantly more effective at satisfying the TLC. Indeed, even the most lateral trajectories remain within the track boundaries.

%---------------------------------
%ORIGINAL VERSION

%This analysis aims to provide empirical validation of the robust trajectories obtained using the methods proposed in this work. For conciseness, the comparison is limited to a nominal trajectory and a trajectory computed using the closed-loop method with only the track limit constraint robustified.
%An LQR controller, designed according to the procedure outlined in Section~\ref{sec:LQR}, is implemented to stabilize the nominal trajectory. In contrast, the robust trajectory is tracked using the controller directly obtained from the optimization process.
